---
title: "univoltine_butterfly_Rmd"
output: html_document
date: "2023-04-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidyr)
library(readr)
library(ggplot2)
library(dplyr)
library(knitr)
library(ggpubr)
```

```{r plot 1}
ggplot(data=univoltine_butterfly,
       aes(forewing_length, rain_jun))+
  geom_boxplot(aes(colour=sex))+
  ggtitle ("The length of male and female butterly wings during the rain in June")+
  labs(y = "Forewing Length", x = "Rain in June")+
  theme(title = element_text(size=7))
```

##R markdown
This plot shows that the length of male and female wings in June is realtivley similar to each other shown by box plot graph displayed. Males have slightly higher 

```{r plot 2}
ggplot(univoltine_butterfly,
       aes(x=jun_mean,
           colour=sex))+
  geom_density(
    alpha=0.7)+
  labs(y = "Density",
       x = "Mean temperature in June",
       title = "Male and female butterflies in different temperatures in June")+
  scale_colour_manual(
    values = c("#117733",
               "#332288",
               "#CC6677"))+
  theme_classic()+
  theme(legend.position = "right")+
  theme(title = element_text(size = 7))
```

##R Markdown
This plot shows that there is a higher density of of females around the temperatures of 14 - 15.5 Celsius. This data shows that at higher temperatures there are more females than males but as the temperature decreases they are relatively similar

```{r plot 3}
ggplot(univoltine_butterfly,
       aes(x = jun_mean,
           y = rain_jun,
           colour = sex))+
  geom_point()+
  labs(x = "Mean June Temperature",
       y = "Rain in June",
       title = "How the mean temperature in June affects rainfall for male and female butterflies")+
  scale_colour_manual(values = c("#CC6677",
                                 "#332288"))+
  theme(title = element_text(size = 6))+
  stat_cor(method="pearson", aes(colour=sex))
```
 ##R Markdown 
 This plot shows the 
  
  

