---
title: "univoltine_butterfly_Rmd"
output: html_document
date: "2023-04-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidyr)
library(readr)
library(ggplot2)
library(dplyr)
library(knitr)
library(ggpubr)
```

```{r plot 1}
ggplot(data=univoltine_butterfly,
       aes(forewing_length, rain_jun))+
  geom_boxplot(aes(colour=sex))+
  ggtitle ("The length of male and female butterly wings during the rain in June")+
  labs(y = "Forewing Length", x = "Rain in June")+
  theme(title = element_text(size=7))
```

##R markdown
This plot shows that the length of male and female wings in June is realtivley similar to each other shown by box plot graph displayed. Males have slightly higher 

```{r plot 2}
ggplot(univoltine_butterfly,
       aes(x=jun_mean,
           colour=sex))+
  geom_density(
    alpha=0.7)+
  labs(y = "Density",
       x = "Mean temperature in June",
       title = "Male and female butterflies in different temperatures in June")+
  scale_colour_manual(
    values = c("#117733",
               "#332288",
               "#CC6677"))+
  theme_classic()+
  theme(legend.position = "right")+
  theme(title = element_text(size = 7))
```

##R Markdown
This plot shows that there is a higher density of of females around the temperatures of 14 - 15.5 Celsius. This data shows that at higher temperatures there are more females than males but as the temperature decreases they are relatively similar

```{r plot 3}
ggplot(univoltine_butterfly,
       aes(x = jun_mean,
           y = rain_jun,
           colour = sex))+
  geom_point()+
  labs(x = "Mean June Temperature",
       y = "Rain in June",
       title = "How the mean temperature in June affects rainfall for male and female butterflies")+
  scale_colour_manual(values = c("#CC6677",                                 "#332288"))+
  theme(title = element_text(size = 6))+
  stat_cor(method="pearson", aes(colour=sex))
```
 ##R Markdown 
This plot shows how the temperature in June affects the rainfall in June for male and female butterflies. Both the R values are negative indicating a negative correlation, meaning the values of one variable increase when the values of the other variable decrease. Both R values are close to zero, females have a closer R value to zero than the males do with an R value of -0.12, compared to the males who's R value is -0.31 meaning females have a weaker linear relationship than the male butterflies do. The p value gives a further understanding as to whether we can conclude that the correlation coefficient is different from zero, and from what we can observe the female butterflies has the furtherst away value from zero. 

```{r plot 4}
ggplot(univoltine_butterfly)+
  geom_smooth(aes(x = jun_mean,
           y = forewing_length,
           colour = sex))+
  labs(y = "Forewing Length",
       x = "Mean temperature in June",
       title = "Identifying how the mean temperature in June affects forewing length in male and female butterflies")+
  scale_colour_manual(
    values = c("#117733",
               "#332288"))+
  theme(title = element_text(size = 4))
``` 
##R Markdowm
This graph shows the relationship between the mean temperature in June and the forewing length of male and female butterflies. Visually the graph shows that as mean temperature increases the forewing length of both male and female butterflies also 
  
  

